---
title: 'MPA/ID Math Camp 2022: Day #5'
output: pdf_document
date: '2022-08-17'
---

### Goals:

1.  Gain additional practice [**reading and manipulating datasets**]{.underline} (`read_csv()` , `dplyr` functions, etc.).
2.  Practice **conditional** statistics (e.g. conditional averages).
3.  Calculate **multi-level** [**group-level statistics**]{.underline} (`group_by()` and `summarise()`).
4.  Practice [**data visualization**]{.underline} with `ggplot2`.

## Continuing International Education Development

Today, we will continue our work on the MICS6 data from yesterday.

#### Question 1

First, reload the `tidyverse` library and read the dataset, called "mics6.csv" in the `data` folder. Assign this data to an object with a name of your choice. For more information on the MICS data, please read the `day_4.Rmd` file from yesterday. Once you've read the data, use one or more functions to list the unique values for the highest grade attained by each child represented in the dataset.

```{r}
library(tidyverse)
df <- read_csv("data/mics6.csv")

unique(df$highest_grade)
```

#### Question 2

In API-209, we will often work with "conditional" statistics. That is, the value of a statistic (e.g. the mean) given that some conditions are met. For example, calculate the average of the `numeracy` variable only among students who completed grade 4 as their highest grade (i.e. students with a value of `highest_grade` that equals 4). This value is the "conditional mean" of `numeracy` for children who completed 4th grade as their highest grade in our sampled countries.

```{r}
fourth <- df %>% filter(highest_grade == 4)
  
mean(fourth$numeracy, na.rm = TRUE)
```

#### Question 3

Calculate the conditional average of `numeracy` for every value of `highest_grade` achieved in our dataset.

**Hint**: You do not need to calculate many different objects and create lots of objects. Review the lesson from yesterday to recall how you might do this with a few `tidyverse` functions.

```{r}
df %>% 
  group_by(highest_grade) %>% 
  summarise(avg_num = mean(numeracy, na.rm = TRUE))
```

#### Question 4

We can also calculate statistics conditional on multiple variables. For example, we might be interested in how growth in numeracy skills across grades differs between countries. `group_by()` can take multiple arguments to create multiple groups. Try editing your code from the previous question to calculate the conditional mean of `numeracy` within each `country` and `highest_grade` separately. Your final dataset should have an average in each country-grade pair (e.g. Ghana 1st grade, Ghana 2nd grade, etc.).

Once you are happy with the results, save this output to an object named `country_grades`.

```{r}
(country_grades <- df %>% 
  group_by(country, highest_grade) %>% 
  summarise(avg_num = mean(numeracy, na.rm = TRUE)))
```

#### Question 5

Let's practice creating new columns in our dataset based on existing values. The orignial `numeracy` column was a binary 1 or 0 for every respondent, so your new average values represent the proportion of respondents who answered yes (1). These values between 0 and 1 are nice, but let's transform them into percentages.

Create a new column in `country_grades` called `num_pct` that multiplies the existing values by 100.

```{r}
country_grades <- country_grades %>% 
  mutate(num_pct = avg_num * 100)
```

#### Question 6

Finally, create a line graph where the highest grade attained is on the x-axis, your `num_pct` numeracy proportion is on the y-axis, and there is one line separately colored for each country. Feel free to consult [Notebook #3](https://colab.research.google.com/drive/11PqVncdckruVj7XOIuhnogMbGtR9jhej?authuser=1#scrollTo=Csp-4ySOrQJP) from the Summer Assignment if you forgot how to set different colors based on values in your dataset.

When your values look right, finish your plot by improving its visual characteristics. At least, please add labels, a title, a caption, and a theme.

```{r}
country_grades %>% 
  ggplot(aes(x = highest_grade, y = num_pct, color = country)) + 
    geom_line() + 
    theme_minimal() +
    labs(x = 'Highest Grade Attained',
         y = 'Average Numeracy',
         title = 'Numeracy Attained by Grade Level and Country',
         caption = 'MICS6 Data')
    
```
