---
title: 'MPA/ID Math Camp 2022: Day #8'
output: pdf_document
date: '2022-08-24'
---

### Goals:

1.  More practice [**creating and customizing tables**]{.underline} in R with `gt`.
2.  Practice with `dplyr` [**data manipulation functions**]{.underline} (e.g. `slice()`).
3.  Creating your own [**functions.**]{.underline}
4.  Data visualization with `ggplot2`.

### **Studying Forcibly Displaced Populations (1964-2008)**

We will continue using the **U.S. Committee for Refugees and Immigrants (USCRI)** dataon refugees and internally displaced persons from yesterday. You can find more information about the dataset [here](https://www.systemicpeace.org/inscrdata.html).

#### Question 1

Load the appropriate libraries and read the `fdp.xls` data into an object called `fdp`.

```{r}

```

#### Question 2

Filter the `fdp` dataset to observations in 1980, and sort (in descending order, from largest to smallest) by the number of refugees hosted in the `host` column. Finally, subset the data to only the top 10 countries.

Once this looks correct to you, save the result to an object.

**Hint:** Remember `slice()` from Summer Assignment [Notebook #2](https://colab.research.google.com/drive/1-Am4YUaSKWEQ8xkpGCRTrIhqQ3RMowQc?authuser=1).

```{r}

```

#### Question 3

Using `select()`, keep only the `country` and `host` columns.

Then, use the remaining columns to make a table with `gt()`. Add a title and subtitle, and [use the tutorial from yesterday](https://gt.rstudio.com/articles/intro-creating-gt-tables.html) to figure out how to make the title bold.

For an extra challenge, try the `cols_label()` function in the tutorial to capitalize the column names and make them bold as well.

```{r}

```

#### Question 4

For a country of your choice, filter the dataset and create a line plot showing the number of refugees hosted for every year in the dataset.

**Hint**: for a cool plotting technique, try adding a `geom_point()` *after* your `geom_line()` to see what happens.

```{r}

```

#### Question 5: Challenge

Adapt your code from Question 4 into a function called `host_plot()` that takes a `country_name` argument. The function should return a plot that, like Question 4, create a line plot showing the number of refugees hosted for every year in the dataset in that country.

Evaluate your function on at least two countries to demonstrate that it runs successfully.

**Hint**: for a title that includes the name of the country, try `paste()`.

```{r}

```
