---
title: 'MPA/ID Math Camp 2022: Day #5'
output: pdf_document
date: '2022-08-17'
---

### Goals:

1.  Gain additional practice [**reading and manipulating datasets**]{.underline} (`read_csv()` , `dplyr` functions, etc.).
2.  Practice **conditional** statistics (e.g. conditional averages).
3.  Calculate **multi-level** [**group-level statistics**]{.underline} (`group_by()` and `summarise()`).
4.  Practice [**data visualization**]{.underline} with `ggplot2`.

## Continuing International Education Development

Today, we will continue our work on the MICS data from yesterday.

#### Question 1

First, reload the `tidyverse` library and read the dataset, called "mics6.csv" in the `data` folder. Assign this data to an object with a name of your choice. For more information on the MICS data, please consult the `day_4.Rmd` file from yesterday. Once you've read the data, use one or more functions to list the number of rows and columns in your dataset.

```{r}

```

#### Question 2

Try finding the average of the `age` column. You may notice that something strange is happening when you try to interpret the value. Investigate all of the values in the `age` column to find out why, and add a comment to your code (with `#`) describing what the problem is.

You may notice that certain "problem" values are included in the data. Create a column with `mutate()` and try to overwrite the values causing the problem you want to change with `ifelse()`. Replace the problem values with `NA`. You can combine knowledge from Summer Assignment Notebooks [#1](https://colab.research.google.com/drive/1dRHlpzfIqyXJd6veU_IMHG7Y93R9DMDb?authuser=1) and [#3](https://colab.research.google.com/drive/11PqVncdckruVj7XOIuhnogMbGtR9jhej?authuser=1#scrollTo=Csp-4ySOrQJP) to try this. Finally, recalculate the average age to see the problem is fixed.

```{r}

```

#### Question 3

Create a histogram using the new `age` values in your data. Feel free to reference the [Visualization Basics](https://rstudio.cloud/learn/primers/3) part of the Summer Assignment and / or to find examples online by searching something like "ggplot histogram."

Open the documentation for the `geom_histogram()` function by running `?geom_histogram` in your console. Read about the `binwidth` argument, and set a value in your function that you like.

Then, go to [this website](https://htmlcolorcodes.com/) (or any other website that allows you to choose a color) and use the colored rectangle to find a color you like. Once you find one, copy the "Hex" code (it will look like a hashtag followed by #F6BA56). Paste that value (including the hashtag) as a character string for the value of the "fill" argument to `geom_histogram()`.

Finally, add labels and a title to your plot, and marvel at its beauty.

```{r}

```

### Let's review answers together

------------------------------------------------------------------------

#### Question 4

In API-209, we will often work with "conditional" statistics. That is, the value of a statistic (e.g. the mean) given that some conditions are met. For example, calculate the average of the `numeracy` variable only among students who completed grade 4 as their highest grade (i.e. students with a value of `highest_grade` that equals 4). This value is the "conditional mean" of `numeracy` for children who completed 4th grade as their highest grade in our sampled countries.

```{r}

```

#### Question 5

We can also calculate statistics conditional on multiple variables. For example, we might be interested in how growth in numeracy skills across grades differs between countries. `group_by()` can take multiple arguments to create multiple groups. Try editing your code from the previous question to calculate the conditional mean of `numeracy` within each `country` and `highest_grade` separately. Your final dataset should have an average in each country-grade pair (e.g. Ghana 1st grade, Ghana 2nd grade, etc.).

Once you are happy with the results, save this output to an object named `country_grades`.

```{r}

```

#### Question 6

Let's practice creating new columns in our dataset based on existing values. The original `numeracy` column was a binary 1 or 0 for every respondent, so your new average values represent the proportion of respondents who answered yes (1). These values between 0 and 1 are conveniently interpreted as proportions, but let's transform them into percentages.

Create a new column in `country_grades` called `num_pct` that multiplies the existing values by 100. This will turn a proportion like 0.52 into a percent like "52."

**Challenge**: if you want to be really fancy, try creating a new column in your dataset called `pct_label` which attaches a `%` sign to the end of your number. How might you do that? You already know the tools you need......

```{r}

```

#### Question 7

Finally, create a line graph where the highest grade attained is on the x-axis, your `num_pct` numeracy proportion is on the y-axis, and there is one line separately colored for each country. Feel free to consult [Notebook #3](https://colab.research.google.com/drive/11PqVncdckruVj7XOIuhnogMbGtR9jhej?authuser=1#scrollTo=Csp-4ySOrQJP) from the Summer Assignment if you forgot how to set different colors based on values in your dataset.

When your values look right, finish your plot by improving its visual characteristics. At least, please add labels, a title, a caption, and a theme.

**Challenge**: if you want to further customize all the colors on this plot (one for each country), try the following: create a vector where the entries are hex codes for individual colors just like above. For example, `my_colors <- c("#ECF0F1", "#5BE877", …` and so on until you have 6 colors -- you need one for every country). Once you have six colors, add the function `scale_color_manual()` to your `ggplot`, with the argument `values` being set equal to the name of your vector. We will learn much more about scale functions soon!

```{r}

```
