---
title: 'MPA/ID Math Camp 2022: Day #5.5 (Extra Practice)'
output: pdf_document
date: '2022-08-19'
---

### Goals:

1.  Gain additional practice [**reading and manipulating datasets**]{.underline} (`read_csv()` , `dplyr` functions, etc.).
2.  Practice [**recoding**]{.underline} and [**creating your own variables.**]{.underline}
3.  Practice advanced [**data visualization**]{.underline} techniques (e.g. function arguments and colors) with `ggplot2`.

## Finishing Study of International Education Development

Today, we will finish our work on the MICS data on children we have used this week. First, again find the average numeracy in our full sample. Then, try finding the average of the `age` column. You may notice that something strange is happening. Investigate all of the values in the `age` column to find out way, and provide a description of why there is an error in the text box below. We will fix the error in the next question.

#### Question 1

```{r}
library(tidyverse)

df <- read_csv("data/mics6.csv")

mean(df$numeracy, na.rm = TRUE)
mean(df$age, na.rm = TRUE)
```

What is the source the problem? Please type your answer in the box below.

```{text}
 
```

#### Question 2

Particular values seem to be causing an error in the age column. Identify what the problem values are, and develop a substantive reason why they might exist with your group. In this question, overwrite the age column and replace those values with a value of NA.

**Hint**: As with many programming tasks, there are multiple ways to do this. However, we recommend creating a column with `mutate()` and trying to overwrite the values you want to change with `ifelse()`. You can combine knowledge from Summer Assignment Notebooks [#1](https://colab.research.google.com/drive/1dRHlpzfIqyXJd6veU_IMHG7Y93R9DMDb?authuser=1) and [#3](https://colab.research.google.com/drive/11PqVncdckruVj7XOIuhnogMbGtR9jhej?authuser=1#scrollTo=Csp-4ySOrQJP) to try this.

```{r}
(df <- df %>% 
  mutate(age = ifelse(age == 9999, NA, age)))
```

#### Question 3

Please recalculate the average and median values of `age` to confirm that your work in Question 2 was done correctly. If so, then the answers you get should look more reasonable. Finally, use the `table()` function to look at all values of the `age` column and confirm they look reasonable. This kind of data exploration and "spot-checking" is extremely important when working with any data, so you know what the values are and the potential problems that might arise.

```{r}

```

#### Question 4

Create a histogram using the new `age` values in your data. Feel free to reference the [Visualization Basics](https://rstudio.cloud/learn/primers/3) part of the Summer Assignment and / or to find examples online by searching something like "ggplot histogram."

Open the documentation for the `geom_histogram()` function by running `?geom_histogram` in your console. Read about the `binwidth` argument, and set a value in your function that you like.

Then, go to [this website](https://htmlcolorcodes.com/) (or any other website that allows you to choose a color) and use the colored rectangle to find a color you like. Once you find one, copy the "Hex" code (it will look like a hashtag followed by #F6BA56). Paste that value (including the hashtag) as a character string for the value of the "fill" argument to `geom_histogram()`.

Finally, add labels and a title to your plot, and marvel at its beauty.

```{r}
df %>% 
  ggplot(aes(x = age)) + 
    geom_histogram(binwidth = 0.25, fill = "#F6BA56") + 
    theme_bw() + 
    labs(title = "Age of Surveyed Children",
         caption = "MICS6 Data",
         x = "Age",
         y = "Number of Students")
```

#### Extra Question

If you finish early, install the `ggthemes` package (in the Console) and choose a nice theme to add to the plot you created in the previous question. Don't forget to load the library in your .Rmd file after you've installed it. You can find a full list of available themes in the package at [this website](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/).

```{r}

```
